---
title: "Aides Drug"
author: "Johana Quintanilla"
date: "Fall 2025"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

I noticed that this is a categorical-categorical type of problem so that means I will use the Cat_Cat-3.rmd helper file to get an idea of how I can make sense out of this.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
females <- filter(triUmp,gender=="F")
```


```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

When observing the bar chart, we came to the realization that 58% of people took the drug and got aids. However, 40% of people that got the placebo, also got aids but this percentage if much smaller than the drug that is supposed to prevent aids. This means that the drug is not doing its job of preventing aids; It gives a higher chance of getting aids.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```

After looking at the numerical results, we saw that it agrees with the graphical results and further proves that the drug does not work properly as it is seen that 58.82% of people got the drug still got aids. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When looking at the inferential results, we came to the conclusion that low P-Values tells us there is no difference between taking the drug and taking the placebo. However, the odds ratio of 1.543902 tells us that individuals are more likely to get aids when taking the drug rather than taking the placebo. In other hands, don't take this drug unless you want aids.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In this assignment we learned how to separate the females from the triUmp chart. We had to analyze the data for only females so we inserted a specific command to split the data for females only. After, we worked our way into analyzing graphical descriptive results, numerical descriptive results and inferential results. In the graphical descriptive results, we learned that the drug is not doing its job of preventing aids; It gives a higher chance of getting aids. In the numerical descriptive results, we learned that it further proves that the drug does not work properly as it is seen that 58.82% of people that got the drug, still got aids. In the inferential results, we learned that we can have a low p-value when it works and when it doesn't work. We just have to check on that. For Instance, low P-Values tells us there is no difference between taking the drug and taking the placebo. However, the odds ratio of 1.543902 tells us that individuals are more likely to get aids when taking the drug rather than taking the placebo.
